# Datashield Demo

## Login and assign

```{r message=FALSE}
# LOAD REQUIRED LIBRARIES
library(datashieldclient)

# LOAD THE TABLE THAT CONTAINS THE LOGIN INFO
load("logindata4bioshare.rda")

# LOGIN TO COLLABORATING SERVERS AND ASSIGN DATA
# the variables to assign
myvar <- list("DIS_CVA","MEDI_LPD","DIS_DIAB","DIS_AMI","GENDER","PM_BMI_CATEGORIAL","LAB_TSC","LAB_HDL","LAB_GLUC_FASTING","PM_BMI_CONTINUOUS")

# run the command to login and assign data
ld <- subset(logindata, server != "chris" & server != "micros" & server != "finrisk")
opals <- datashield.login(logins=ld, assign=TRUE, variables=myvar)
```

Get the total number of participants:

```{r}
sum(unlist(datashield.aggregate(opals,"length(D$GENDER)")))
```

## Univariate distribution

Tabulate some binary variables:

```{r message=FALSE}
ds.table1d(datasources=opals, xvect=quote(D$DIS_CVA))
```

```{r message=FALSE}
ds.table1d(datasources=opals, xvect=quote(D$MEDI_LPD))
```

```{r message=FALSE}
ds.table1d(datasources=opals, xvect=quote(D$DIS_DIAB))
```

```{r message=FALSE}
ds.table1d(datasources=opals, xvect=quote(D$DIS_AMI)) 
```

```{r message=FALSE}
ds.table1d(datasources=opals, xvect=quote(D$GENDER))
```

Now see what happens if attempt to tabulate a continuous variable:

```{r}
ds.table1d(datasources=opals, xvect=quote(D$PM_BMI_CONTINUOUS))
```

## Quantile Mean

Display quantile (you can choose to display the combine (default) or separate quantiles)

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$LAB_TSC))
```

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$LAB_TSC), type="split")
```

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$LAB_HDL))
```

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$LAB_HDL), type="split")
```

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$PM_BMI_CONTINUOUS), type="split")
```

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$LAB_GLUC_FASTING))
```

## Histogram

Plot histograms:

```{r message=FALSE}
ds.histogram(datasources=opals, xvect=quote(D$LAB_HDL))
```

```{r message=FALSE}
ds.histogram(datasources=opals, xvect=quote(D$LAB_HDL), type="split")
```

```{r message=FALSE}
ds.histogram(datasources=opals, xvect=quote(D$PM_BMI_CONTINUOUS))
```

```{r message=FALSE}
ds.histogram(datasources=opals, xvect=quote(D$PM_BMI_CONTINUOUS), type="split")
```

## Questions

### Question 1

To see if there is a difference across two variables we can use t.test:

```{r message=FALSE}
ds.t.test(datasources=opals, x=quote(D$LAB_TSC), y=quote(D$LAB_HDL))
```

```{r message=FALSE}
ds.t.test(datasources=opals, x=quote(D$LAB_TSC), y=quote(D$LAB_HDL), type="split")
```

We can save the outcome to our workspace by adding in '<-':

```{r message=FALSE}
q11 <- ds.t.test(datasources=opals, x=quote(D$LAB_TSC), y=quote(D$LAB_HDL))
```

### Question 2

To explore the relationship between 'LAB_GLUC_FASTING' and 'PM_BMI_CATEGORIAL' let us tabulate the factor variable

```{r message=FALSE}
ds.table1d(datasources=opals, xvect=quote(D$PM_BMI_CATEGORIAL))
```

Use the glm function the explore the relationship:

```{r message=FALSE}
glm.mod2 <- ds.glm(datasources=opals, formula=D$LAB_GLUC_FASTING ~ D$PM_BMI_CATEGORIAL, family=quote(gaussian), maxit=quote(20))
# display a summary of the results
glm.mod2
```

### Question 3

To determine a trend over two categorical factor variables let us cross tabulate those two variables using the function 'ds.table2d'

```{r message=FALSE}
ds.table2d(datasources=opals, xvect=quote(D$DIS_DIAB), yvect=quote(D$PM_BMI_CATEGORIAL))
```

We can formally test the differing proportions across categories by using the function 'ds.table2' and look into the chi-squared results

```{r message=FALSE}
q3 <- ds.table2d(datasources=opals, xvect=quote(D$DIS_DIAB), yvect=quote(D$PM_BMI_CATEGORIAL))
q3$CHI2.TESTS.FOR.HOMOGENEITY
```

### Question 4

To determine if total serum cholesterol levels are associated with HDL cholesterol levels. For example, this can be carried out by plotting a scatter plot or performing a correlation or using linear regression using LAB_TSC as an outcome and LAB_HDL as a covariate other covariates can be added to determine other significant variables correlated with total serum cholesterol.

```{r message=FALSE}
ds.heatmapplot(datasources=opals, xvect=quote(D$LAB_TSC), yvect=quote(D$LAB_HDL))
```

Now let us produce the hetamap plots of the studies separately:

```{r message=FALSE}
ds.heatmapplot(datasources=opals, xvect=quote(D$LAB_TSC), yvect=quote(D$LAB_HDL), type="split")
```

Generate pooled contour plot:

```{r message=FALSE}
ds.contourplot(datasources=opals, xvect=quote(D$LAB_TSC), yvect=quote(D$LAB_HDL))
```

```{r message=FALSE}
glm.mod4 <- ds.glm(datasources=opals, formula=D$LAB_TSC~D$LAB_HDL, family=quote(gaussian), maxit=quote(20))
glm.mod4
```


### Question 5

To determine predictors of binary outcomes we need to use the function glm. The glm function can be used to analyse any input or output whether it be categorical or numerical. 
* If the outcome to analyse is numeric then we use the “gaussian” option of the glm.  
* If the outcome contains two categories we can use the “binomial” option.  

So the type of glm function we need to answer research question 5 is the one which has family input “binomial”.  

First let us tabulate the two variables:

```{r message=FALSE}
ds.table2d(datasources=opals, xvect=quote(D$DIS_DIAB), yvect=quote(D$PM_BMI_CATEGORIAL))
```

Now run glm to predict diabetes status using categorical bmi:

```{r message=FALSE}
glm.mod5_1 <- ds.glm(datasources=opals, formula=D$DIS_DIAB ~ D$PM_BMI_CATEGORIAL, family=quote(binomial), maxit=quote(20))
glm.mod5_1
```

Run glm to predict diabestes using gender, continuous bmi and hdl cholesterol;

```{r message=FALSE}
glm.mod5_2 <- ds.glm(datasources=opals, formula=D$DIS_DIAB~D$GENDER+D$PM_BMI_CONTINUOUS+D$LAB_HDL, family=quote(binomial), maxit=quote(20))
glm.mod5_2
```


### Question 6

To determine predictors of taking lipid reducing medications. For example, a logistic regression can, again, be used to determine whether some HOP variables are significantly associated with a taking lipid reducing medications, MEDI_LPD.

```{r message=FALSE}
glm.mod6_1 <- ds.glm(datasources=opals, formula=D$MEDI_LPD~D$GENDER*D$LAB_HDL+D$PM_BMI_CONTINUOUS+D$LAB_TSC, family=quote(binomial), maxit=quote(20))
glm.mod6_1
```

Quantile of the variable 'LAB_HDL':

```{r message=FALSE}
ds.quantilemean(datasources=opals, xvect=quote(D$LAB_HDL))
```

Substract the mean:

```{r message=FALSE}
datashield.assign(opals, 'HDL.1.5', quote(D$LAB_HDL-1.5))
```

Run another glm analysis using the above adjusted hdl variable:

```{r message=FALSE}
glm.mod6_2 <- ds.glm(datasources=opals, formula=D$MEDI_LPD~D$GENDER*HDL.1.5+D$PM_BMI_CONTINUOUS+D$LAB_TSC, family=quote(binomial), maxit=quote(20))
glm.mod6_2
```

## Logout

```{r message=FALSE,results='hide'}
datashield.logout(opals)
```
